---
output:
  md_document:
    variant: gfm
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-",
  eval = FALSE
)
```

# cnf

Tools to manage configuration files in R projects.

## Examples

Assume you have a configuration file located at `~/.config.R`. It may look
something like:

**~/.cnf.R**

```{r}
user = "user"
password = "pwd"
...
```

Now the question is, how do you make these settings available to your R project.
`cnf` provides the following mechanisms:

- `register`: Register a configuration file. It adds to a list of configurations.
- `getcnf`: get the value of a config. It is always a `list` returned with the
values in the config as elements.

```{r}
cnf::register(config = "~/.config.R")
cnf::getcnf("config")
```

With `get` we get the config for a given name. You can and should give project
specific names to rule out name clashes between them:

```{r}
cnf::register(
  projA = "pathA/config.R",
  projB = "pathB/config.R"
)
```

We now have access to these configs using:

```{r}
cnf::getcnf("projA")
cnf::getcnf("projB")
```

Note that we can override configurations using this approach; simply by using
the same name twice. A warning is raised, however this may be entirely by
intention. Consider that you have projects A and B. A has its own configuration and
loads it. B depends on A but needs to reconfigure project A. Think of the number
of cores, database credentials, and the like. Using this approach we will
inherit the configuration but can override what needs to be overridden.

To make this work the order of registering is important. Thus the registration
goes into the `.onLoad` hook of your package. If you do not have a package: you
are doomed, go and write a package!

```{r}
.onLoad <- function(libname, pkgname) {
  cnf::register(
    projA = "pathA/config.R",
    projB = "pathB/config.R",
    maybe = TRUE,
    quiet = TRUE,
    warn = FALSE
  )
}
```

- `maybe` allows this call to fail, but will print a message. This is important
when we install a package and have no configuration, yet.
- `quiet` suppresses all warnings and messages.
- `warn` turns warnings into messages.

Sometimes configurations are stored as R object within the package. As R code
and in version control. We can register them preferably as load hook:

```{r}
.onLoad <- function(libname, pkgname) {
  cnf::register(
    cnf = configObject
  )
}
```

In this case `configObject` is a list. We can go wild and provide a
fallback/default configuration and allow for override by a file or environment
variable:

```{r}
.onLoad <- function(libname, pkgname) {
  cnf::register(
    config = configObject,
    config = "~/.config.R",
    config = Sys.getenv("CONFIG"),
    maybe = TRUE
  )
}
```

Using this pattern, we can override the configuration, or parts of it, by an
environment variable or config file. The file path in the environment variable
has highest priority.

Happy coding...
