---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-",
  eval = FALSE
)
```

# config

Tools to manage configuration files for R projects.

## Examples

Assume you have a configuration file located at `~/.config.R`. It may look
something like:

**~/.config.R**

```{r}
user <- "user"
password <- "pwd"
...
```

Now the question is, how do you make these settings available to your R project.
`config` provides the following mechanisms:

- `register`: Register a configuration file. It adds to a list of configurations.
- `get`: get the value of a config. It is always a `list` returned with the
values in the config as elements.
- `store`: a *store* of all registered configurations.

```{r}
config::register("~/.config.R")
config::get("config")
config::store$config(user = "test-user")
```

As we can see, we have two options to access values in the configuration: `get`
and `store`. `get` is easy to understand: we get the config for a given name.
`store` is an environment. When we register a new configuration, a function is
added to the store to 'generate' the config. Using this function we can override
or add elements in the config, or just retrieve the values as they are. This can
be useful in development and debugging sessions.

One more thing you may have recognised: The name *config* has been deduced from
the filepath. By default we use the filename provided without the file extension
and after stripping the leading dot. If you name all your configuration files
`~/.project/config.R`, this is a bit ambiguous. Instead you can and should give
project specific names:

```{r}
config::register(
  projA = "pathA/config.R",
  projB = "pathB/config.R"
)
```

We now have access to these configs using:

```{r}
config::get("projA")
config::get("projB")
config::store$projA()
config::store$projB()
```

Note that we can override configurations using this approach; simply by using
the same name twice. A warning is raised, however this may be entirely by
intention. Consider that you have projects A and B. A has its own config and
loads it. B depends on A but needs to reconfigure project A. Think of the number
of cores, database credentials, and the like. Using this approach we will
inherit the configuration but can override what needs to be overridden.

To make this work the order of registering is important. Thus the registration
goes into the `.onLoad` hook of your package. If you do not have a package: you
are doomed, go and write a package!

```{r}
.onLoad <- function(libname, pkgname) {
  config::register(
    projA = "pathA/config.R",
    projB = "pathB/config.R",
    maybe = TRUE,
    quiet = TRUE,
    warn = FALSE
  )
}
```

- `maybe` allows this call to fail, but will print a message. This is important
when we install a package and have no configuration, yet.
- `quiet` suppresses all warnings and messages.
- `warn` turns warnings into messages.

For some reason we sometimes have configuration in our packages. As R code and
in version control. We can register them as configuration, preferably as load
hook:

```{r}
.onLoad <- function(libname, pkgname) {
  config::register(
    config = configObject
  )
}
```

In this case `configObject` is a list. We can go wild and provide a
fallback/default configuration and allow for override by a file or environment
variable:

```{r}
.onLoad <- function(libname, pkgname) {
  config::register(
    config = configObject,
    config = "~/.config.R",
    config = Sys.getenv("CONFIG"),
    maybe = TRUE
  )
}
```

Using this pattern, we can override the configuration, or parts of it, by an
environment variable or config file. But it is optional. The file path in the
environment variable has highest priority.

Happy coding...
